# (PART\*) Appendix {-} 

# FHSWF Package

Für die Bearbeitung im Rahmen des Kurses steht Ihnen das `fhswf` Package zur Verfügung. Dieses ergänzt R um einige spezielle Funktionen, stellt weitere Datensätze bereit und installiert Vorlagen (Templates) für Seminar-, Gruppen- und Abschlussarbeiten sowie für Präsentationenen. 

Das `fhswf` Package befindet sich aktuell noch in der Entwicklung und ist daher noch nicht im zentralen offiziellen Package Archiv (CRAN) gelistet. Für die Installation benötigen Sie eine Funktion aus dem Paket `remotes`, dass einmalig installiert werden muss. Die jeweils neuste `fhswf`-Paketversion erhalten Sie mit dem folgenden Programmcode.

```{r, eval = FALSE}
install.packages("remotes")              # Muss nur einmalig ausgeführt werden
remotes::install_github("bchwtz/fhswf")  # Neueste Paketversion installieren 
```

# Daten 2022

## NYC Airbnb

```{r, eval = FALSE}
library(fhswf)       # Laden des fhswf-Packages
View(nycairbnb)      # Anzeigen der Daten
nycairbnb            # Anzeigen der Daten auf der Konsole
?nycairbnb           # Anzeigen der Datenbeschreibung
```

### Beispiel

```{r}
summary(nycairbnb$price) # Deskriptive Statistik des Preises pro Übernachtung
```

## Autopreise

```{r, eval = FALSE}
library(fhswf)       # Laden des fhswf-Packages
View(carprices)      # Anzeigen der Daten
carprices            # Anzeigen der Daten auf der Konsole
?carprices           # Anzeigen der Datenbeschreibung
```

### Beispiel

```{r}
summary(carprices$price) # Deskriptive Statistik des Kaufpreises
```


## 2022

### Autos

https://www.kaggle.com/code/dronax/car-prices-dataset/data

### NYC Airbnb

https://www.kaggle.com/datasets/dgomonov/new-york-city-airbnb-open-data

### WHO Life Expectancy

https://www.kaggle.com/datasets/kumarajarshi/life-expectancy-who

## 2021

### Taxi Trips

#### Vorbereitungen

```{r, eval = FALSE, code=xfun::read_utf8("data/2021_fhswf_stat_taxi.R")}

```

#### Datenbeschreibung

**Variablen**

-   `id`: a unique identifier for each trip
-   `vendor_id`: a code indicating the provider associated with the trip record
-   `pickup_datetime`: date and time when the meter was engaged
-   `dropoff_datetime`: date and time when the meter was disengaged
-   `passenger_count`: the number of passengers in the vehicle (driver entered value)
-   `pickup_longitude`: the longitude where the meter was engaged
-   `pickup_latitude`: the latitude where the meter was engaged
-   `dropoff_longitude`: the longitude where the meter was disengaged
-   `dropoff_latitude`: the latitude where the meter was disengaged
-   `store_and_fwd_flag`: This flag indicates whether the trip record was held in vehicle memory before sending to the vendor because the vehicle did not have a connection to the server - Y=store and forward; N=not a store and forward trip
-   `trip_duration`: duration of the trip in seconds

**Ergänzte Variablen**

-   `osrm_distance`: the travel distance by car in kilometers as calculated by the Open Streetmap Routing Machine
-   `osrm_duration`: the estimated duration for travelling the route by car in minutes as estimated by the Open Streetmap Routing Machine
-   `geodesic_distance`: the geodesic distance in meters between the pickup and dropoff coordinates calculated using the method of Karney (2013)
-   `weekday:` the day of the week when the trip was started

#### Daten laden

```{r}
# load the data
x <- read.csv("https://bchwtz.github.io/bchwtz-stat/data/2021_fhswf_stat_taxi.csv")
```

#### Tips & Tricks

```{r}
# working with dates and times
x$pickup_datetime <- as.POSIXlt(x$pickup_datetime, tz="America/New_York")
x$pickup_datetime[1]
weekdays(x$pickup_datetime[1])
months(x$pickup_datetime[1])
```

#### Quellen

-   <https://www.kaggle.com/c/nyc-taxi-trip-duration/>
-   <https://www1.nyc.gov/site/tlc/about/tlc-trip-record-data.page>

### Hauspreise

#### Vorbereitungen

```{r, eval = FALSE, code=xfun::read_utf8("data/2021_fhswf_stat_housing.R")}

```

#### Datenbeschreibung

-   `id`: a unique identifier for each offer
-   `price`: the requested price in Euro to buy the building
-   `type`: the type of the offered building
-   `living_space`: the size of the usable living space in the building in square meters
-   `lot_size`: the size of the associated lot in square meters
-   `additional_area`: the size of the additionally available area in square meters that is not qualified living space due to legal reasons
-   `rooms`: the total number of rooms in the building
-   `bedrooms`: the number of bedrooms in the building
-   `bathrooms`: the number of bathrooms in the building
-   `floors`: number of storys in the building
-   `year_construction`: the year in which the building was first constructed
-   `year_modernization`: the year in which the building was lastly renovated
-   `facility_quality`: an indication of quality of the installed fixtures, facilities and if applicable furniture
-   `condition`: the general condition of the building and facilities
-   `heating`: the primary technology for heating the building
-   `energy_source`: the primary source of energy for heating and warm water
-   `energy_certificate`: an indication wheter and energy consumption certificate is required and/or available
-   `energy_certificate_type`: the type of the available energy certification
-   `energy_consumption`: the required amount of energy in $kWh/m^2$ per year to heat the building
-   `energy_efficiency_class`: an indication of the quality of the thermal insulation of the building according to the german EnEV, ranging from A+ (best) to H (worst)
-   `state`: the federal state of germany in which the building is located
-   `district`: the respective district of the federal stat in which the building is located
-   `city`: the name of the city to whom the building is allocated
-   `parking_lot`: the amount of registered parking lots that belong to the building
-   `parking_lot_type`: the type of the available parking lots

#### Daten laden

```{r, eval=FALSE}
# load the data
x <- read.csv("https://bchwtz.github.io/bchwtz-stat/data/2021_fhswf_stat_housing.csv")
```

#### Quellen

-   <https://www.kaggle.com/scriptsultan/german-house-prices>

### Human Ressources

#### Vorbereitungen

```{r, eval = FALSE, code=xfun::read_utf8("data/2021_fhswf_stat_hr.R")}

```

#### Datenbeschreibung

-   `EmployeeName`: Employee's full name d
-   `EmpID`: Employee ID is unique to each employee\
-   `MarriedID`: Is the person married (1 or 0 for yes or no)\
-   `MaritalStatusID`: Marital status code that matches the text field MaritalDesc
-   `EmpStatusID`: Employment status code that matches text field EmploymentStatus
-   `DeptID   Department`: ID code that matches the department the employee works in\
-   `PerfScoreID`: Performance Score code that matches the employee's most recent performance score\
-   `FromDiversityJobFairID`: Was the employee sourced from the Diversity job fair? 1 or 0 for yes or no\
-   `PayRate`: The person's hourly pay rate. All salaries are converted to hourly pay rate
-   `Termd`: Has this employee been terminated - 1 or 0\
-   `PositionID`: An integer indicating the person's position
-   `Position`: The text name/title of the position the person has\
-   `State`: The state that the person lives in\
-   `Zip`: The zip code for the employee\
-   `DOB`: Date of Birth for the employee\
-   `Sex`: Sex - M or F\
-   `MaritalDesc`: The marital status of the person (divorced, single, widowed, separated, etc)\
-   `CitizenDesc`: Label for whether the person is a Citizen or Eligible NonCitizen\
-   `HispanicLatino`: Yes or No field for whether the employee is Hispanic/Latino
-   `RaceDesc`: Description/text of the race the person identifies with
-   `DateofHire`: Date the person was hired\
-   `DateofTermination`: Date the person was terminated, only populated if, in fact, Termd = 1\
-   `TermReason`: A text reason / description for why the person was terminated\
-   `EmploymentStatus`: A description/category of the person's employment status. Anyone currently working full time = Active\
-   `Department`: Name of the department that the person works in
-   `ManagerName`: The name of the person's immediate manager\
-   `ManagerID`: A unique identifier for each manager
-   `RecruitmentSource`: The name of the recruitment source where the employee was recruited from\
-   `PerformanceScore`: Performance Score text/category (Fully Meets, Partially Meets, PIP, Exceeds)\
-   `EngagementSurvey`: Results from the last engagement survey, managed by our external partner\
-   `EmpSatisfaction`: A basic satisfaction score between 1 and 5, as reported on a recent employee satisfaction survey
-   `SpecialProjectsCount`: The number of special projects that the employee worked on during the last 6 months
-   `LastPerformanceReviewDate`: The most recent date of the person's last performance review\
-   `DaysLateLast30`: The number of times that the employee was late to work during the last 30 days\
-   `Absences`: The number of times the employee was absent from work

#### Daten laden

```{r,}
# load the data
x <- read.csv("https://bchwtz.github.io/bchwtz-stat/data/2021_fhswf_stat_hr.csv")
```

#### Tips & Tricks

```{r}
# working with dates and times
x$Date_Hiring <- as.Date(x$DateofHire,format='%m/%d/%Y')
x$Date_Termination <- as.Date(x$DateofTermination,format='%m/%d/%Y')

# example difference between two timepoints
difftime(x$Date_Termination[2],x$Date_Hiring[2], units = "days")
```

#### Quellen

-   <https://www.kaggle.com/rhuebner/human-resources-data-set>
-   <https://rpubs.com/rhuebner/hr_codebook_v14>
